Temporary Trade Barriers: \\ When Will They End?

### 2017-10-31 presentation.tex
update presentation for jr seminar

\pause
The probability that an AD duty gets renewed
\pause
\begin{itemize}[<+->]
	\item Increases in lobbying effort
	\item Decreases in the MFN tariff
	\item Is invariant to the trading partner's tariff
	\item Increases in the profitability of the import-competing sector
	\item Increases in the strength of the lobby
	\item May be concave in the AD duty
\end{itemize}
\end{frame}

@@

\begin{frame}{Political Structure}
In Home country (foreign is passive):
\pause
\begin{itemize}[<+->]
	\item Dept. of Commerce / Int'l Trade Commission
		\begin{itemize}[<+->]
			\item Can renew AD duties
			\item Susceptible to influence of lobbying, perhaps both direct and indirect
			\item Modeled in reduced form
		\end{itemize}
%\pause
	\item A Single Lobby
		\begin{itemize}
			\item Represents import-competing sector, $X$
		\end{itemize}	
\end{itemize}

@@

\pause
\vskip.1in
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Assumption 1}
\begin{enumerate}
  \pause
	\item $\ga(e,\ta)$ is increasing and concave in $e$ for all $\ta \in \Theta$.
\end{enumerate}
\end{beamerboxesrounded}
\end{frame}

@@

\vskip.1in
\pause
\begin{itemize}[<+->]
	\item $e$: Lobbying effort
	\item $\tau^a$: home import tariff under trade agreement
	\item $\tau^{\textit{ad}}$: home import tariff equivalent under anti-dumping duties
\end{itemize}

@@

\section{Uncertainty}
\subsection{}
\begin{frame}{Timeline}
\begin{enumerate}[<+->]
	\item \textbf{Import-competing firms lobby DOC/ITC to renew AD duties}
	\item {\color{gray} \textbf{Uncertainty is resolved}}
	\item \textbf{DOC/ITC decide whether to renew duties}
	\item {\color{gray} Private actors make production, consumption decisions}
\end{enumerate}

@@

@@

@@

\begin{frame}{Government}
  $G$ renews AD duties if its utility is higher under AD duties than trade agreement tariff
	\pause
  \begin{itemize}
		\item Preferences are ex-ante uncertain through $\ta$
		\pause
		\item When does $G$ renew AD duties? \\
	\pause
  \vskip.1in
    $b(e,\tau^a,\tau^{\textit{ad}})$: probability $G$ prefers $\tau^{ad}$ to $\tau^a$ for a given effort level $e$
\end{itemize}

\pause
\vskip.25in
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Lemma 1}
  The probability that $G$ renews AD duties is increasing and concave in lobbying effort $e$ $\left(\text{i.e. } \frac{\partial b}{\partial e} \geq 0, \ \frac{\partial^2 b}{\partial e^2} \leq 0  \right)$.
\end{beamerboxesrounded}

@@

\vskip.25in
\pause
There's both a direct effect and an indirect effect through lobby's incentives, and both are negative:
\[
  \frac{\partial b}{\partial e} \frac{\partial e}{\partial \tau^a} + \frac{\partial b}{\partial \tau^a}
\]
\end{frame}

@@

Assuming trading partner does not retaliate
\pause
\begin{itemize}
	\item No difference in foreign tariff under AD duty and $\tau^a$. So no effect on $G$'s incentives (either direct or indirect)
\end{itemize}

@@

@@

\pause
\textit{NOTE: this is not quite right, but some version of it will be} \\
Assume $\pi(\cdot)$ shifts up uniformly for all $\tau$.
\pause
\begin{itemize}[<+->]
	\item Convexity of profits $\Rightarrow$ $G$'s marginal benefit of providing protection goes up
	\item Convexity of profits $\Rightarrow$ return from lobbying increases
\end{itemize}

\vskip.25in
\pause
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Result 3}
  The total probability that $G$ renews AD duties is increasing in the profitability of the import-competing sector.
\end{beamerboxesrounded}
\end{frame}

\pause
Assume $\ga(\cdot,\cdot)$ shifts up uniformly for all $(e,\ta)$ pairs.
\pause
\begin{itemize}[<+->]
	\item $G$ gives more weight to firms' benefit
	\item Lobbying incentives are unchanged
\end{itemize}

\vskip.25in
\pause
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Result 3}
  The total probability that $G$ renews AD duties increases when the weighting function shifts up exogenously and uniformly.
\end{beamerboxesrounded}
\end{frame}

\begin{frame}{Protection from AD Duties}

\pause
When $\tau^{\textit{ad}}$ increases, two effects on $G$'s incentives:
\pause
\begin{itemize}[<+->]
	\item Social welfare decreases, pushes for decrease in renewal probability 
	\item (Over-weighted) import-competing profits increase, pushes for increase in renewal probability 
\end{itemize}

\vskip.2in
\pause
Indirect effect is of same sign as direct effect
\pause
\begin{itemize}[<+->]
	\item When $\tau^{\textit{ad}}$ (i.e. close to social optimum), second effect dominates $\Rightarrow$ increase in renewal probability
	\item Effect may be concave
\end{itemize}

@@

\begin{frame}{Feedback, Comments from InsTED presentation}
\begin{itemize}[<+->]
	%\item Could leverage 5-year renewal cycle vs. 4-year re-election cycle (Sarah Danzman, Indiana U)
	\item Look for Kara Reynold's ``Political Economy of Antidumping Reviews (Shushanik)
	\item Need to have first stage (Jee-Hyeong, Gary Lyn)
		\begin{itemize}
			\item Look at Park and Blonigen (Jee-Hyeong)
			\item Cost differential for firms to get initial duty vs. renewal?
			\item Some firms choose not to pursue renewal when DOC contacts them 30-days before (Sasha)
		\end{itemize}
	\item Lobbying effort may be endogenous: if prob. of success is lower, maybe it's not worth providing effort (Mostafa)
	\item Uncertainty could be, in part, about how mobile labor is (Kishore)
\end{itemize}

>>>>

\pause
The probability that AD duties gets renewed
\pause
\begin{itemize}[<+->]
	\item increases in lobbying effort
	\item decreases in the MFN tariff
	\item is invariant to the trading partner's tariff
	\item increases in the profitability of the import-competing sector
	\item increases in the strength of the lobby
	\item AD duties: complicated
\end{itemize}
\end{frame}

@@

\begin{frame}{Political Structure}
In Home country (foreign is passive):
\pause
\begin{itemize}[<+->]
	\item Dept. of Commerce / Int'l Trade Commission ($G$)
		\begin{itemize}[<+->]
			\item Can renew AD duties
			\item Susceptible to influence of lobbying, perhaps both direct and indirect
			\item Modeled in reduced form
		\end{itemize}
%\pause
	\item A Single Lobby
		\begin{itemize}
			\item Represents import-competing sector, $X$
		\end{itemize}	
\end{itemize}

@@

\pause
\vskip.2in
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Assumption 1}
\begin{enumerate}
  \pause
	\item $\ga(e,\ta)$ is increasing and concave in $e$ for all $\ta \in \Theta$.
\end{enumerate}
\end{beamerboxesrounded}
\end{frame}

@@

\vskip.1in
\pause
\begin{itemize}[<+->]
	\item $\tau^a$: home import tariff under trade agreement
	\item $\tau^{\textit{ad}}$: home import tariff equivalent under anti-dumping duties
\end{itemize}

@@

\section{Uncertainty}
\subsection{}

\begin{comment}
\begin{frame}{Timeline}
\begin{enumerate}[<+->]
	\item \textbf{Import-competing firms lobby DOC/ITC to renew AD duties}
	\item {\color{gray} \textbf{Uncertainty is resolved}}
	\item \textbf{DOC/ITC decide whether to renew duties}
	\item {\color{gray} Private actors make production, consumption decisions}
\end{enumerate}

@@

\end{frame}
\end{comment}

@@

\pause
\vskip.2in
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Assumption 2}
\begin{enumerate}
  \pause
	\item $\ta$ is uniformly distributed on some $\Theta=\left[\underline{\ta},\overline{\ta}\right]$.
\end{enumerate}
\end{beamerboxesrounded}

@@

\begin{frame}{Government}
  $G$ renews AD duties if its utility is higher under AD duties than trade agreement tariff
	\pause
  \begin{itemize}
		\item Preferences are ex-ante uncertain through $\ta$
		\pause
		\item When does $G$ renew AD duties? \\
	\pause
  \vskip.1in
    $r(e,\tau^a,\tau^{\textit{ad}})$: probability $G$ prefers $\tau^{ad}$ to $\tau^a$ for a given effort level $e$
\end{itemize}

\pause
\vskip.25in
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Lemma 3}
  The probability that $G$ renews AD duties is increasing and concave in lobbying effort $e$ $\left(\text{i.e. } \frac{\partial r}{\partial e} \geq 0, \ \frac{\partial^2 r}{\partial e^2} \leq 0  \right)$.
\end{beamerboxesrounded}

@@

\vskip.25in
\pause
There's both a direct effect and an indirect effect through lobby's incentives, and both are negative:
\[
  \frac{\partial r}{\partial e} \frac{\partial e}{\partial \tau^a} + \frac{\partial r}{\partial \tau^a}
\]
\end{frame}

@@

Assuming trading partner does not retaliate
\pause
\begin{itemize}
	\item No difference in foreign tariff under AD duties and $\tau^a$. So no effect on $G$'s incentives (either direct or indirect)
\end{itemize}

@@

\begin{frame}{Protection through AD Duties}

\pause
When $\tau^{\textit{ad}}$ increases, direct effect is negative: social welfare decreases, so $G$ less likely to renew

\vskip.2in
\pause
Indirect effect is positive: higher $\tau^{\textit{ad}}$ increases incentive to exert lobbying effort

\vskip.2in
\pause
Magnitudes are not directly comparable
\begin{itemize}
	\item Working to endogenize $\tau^{\textit{ad}}$
\end{itemize}

\pause
Assume $\ga(\cdot,\cdot)$ increases weakly for all $(e,\ta)$ pairs.
\pause
\begin{itemize}[<+->]
	\item $G$ gives more weight to firms' benefit
	\item Lobbying incentives are unchanged
\end{itemize}

\vskip.25in
\pause
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Result 4}
  The total probability that $G$ renews AD duties increases when the weighting function shifts up exogenously.
\end{beamerboxesrounded}
\end{frame}

@@

\pause
Both lobby and $G$ care on the margin about the difference between profits under $\tau^{\textit{ad}}$ and $\tau^a$
\pause
\begin{itemize}[<+->]
	\item Shifts in the profit function matter to the extent they change this difference 
\end{itemize}

\vskip.25in
\pause
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Result 5}
  The total probability that $G$ renews AD duties increases (decreases) when there is an increase (decrease) in the gap between profits under $\tau^{\textit{ad}}$ and profits under $\tau^a$.
\end{beamerboxesrounded}
\end{frame}

@@

\begin{comment}
\begin{frame}{Feedback, Comments from InsTED presentation}
\begin{itemize}[<+->]
	%\item Could leverage 5-year renewal cycle vs. 4-year re-election cycle (Sarah Danzman, Indiana U)
	\item Look for Kara Reynold's ``Political Economy of Antidumping Reviews (Shushanik)
	\item Need to have first stage (Jee-Hyeong, Gary Lyn)
		\begin{itemize}
			\item Look at Park and Blonigen (Jee-Hyeong)
			\item Cost differential for firms to get initial duties vs. renewal?
			\item Some firms choose not to pursue renewal when DOC contacts them 30-days before (Sasha)
		\end{itemize}
	\item Lobbying effort may be endogenous: if prob. of success is lower, maybe it's not worth providing effort (Mostafa)
	\item Uncertainty could be, in part, about how mobile labor is (Kishore)
\end{itemize}

\end{frame}
\end{comment}
